\documentclass[12pt,letterpaper]{article}
\usepackage{graphicx,textcomp}
\usepackage{natbib}
\usepackage{setspace}
\usepackage{fullpage}
\usepackage{color}
\usepackage[reqno]{amsmath}
\usepackage{amsthm}
\usepackage{fancyvrb}
\usepackage{amssymb,enumerate}
\usepackage[all]{xy}
\usepackage{endnotes}
\usepackage{lscape}
\newtheorem{com}{Comment}
\usepackage{float}
\usepackage{hyperref}
\newtheorem{lem} {Lemma}
\newtheorem{prop}{Proposition}
\newtheorem{thm}{Theorem}
\newtheorem{defn}{Definition}
\newtheorem{cor}{Corollary}
\newtheorem{obs}{Observation}
\usepackage[compact]{titlesec}
\usepackage{dcolumn}
\usepackage{tikz}
\usetikzlibrary{arrows}
\usepackage{multirow}
\usepackage{xcolor}
\newcolumntype{.}{D{.}{.}{-1}}
\newcolumntype{d}[1]{D{.}{.}{#1}}
\definecolor{light-gray}{gray}{0.65}
\usepackage{url}
\usepackage{listings}
\usepackage{color}

\definecolor{codegreen}{rgb}{0,0.6,0}
\definecolor{codegray}{rgb}{0.5,0.5,0.5}
\definecolor{codepurple}{rgb}{0.58,0,0.82}
\definecolor{backcolour}{rgb}{0.95,0.95,0.92}

\lstdefinestyle{mystyle}{
	backgroundcolor=\color{backcolour},   
	commentstyle=\color{codegreen},
	keywordstyle=\color{magenta},
	numberstyle=\tiny\color{codegray},
	stringstyle=\color{codepurple},
	basicstyle=\footnotesize,
	breakatwhitespace=false,         
	breaklines=true,                 
	captionpos=b,                    
	keepspaces=true,                 
	numbers=left,                    
	numbersep=5pt,                  
	showspaces=false,                
	showstringspaces=false,
	showtabs=false,                  
	tabsize=2
}
\lstset{style=mystyle}
\newcommand{\Sref}[1]{Section~\ref{#1}}
\newtheorem{hyp}{Hypothesis}

\title{Problem Set 2}
\date{Due: February 18, 2024}
\author{Applied Stats II}


\begin{document}
	\maketitle
	\section*{Instructions}
	\begin{itemize}
		\item Please show your work! You may lose points by simply writing in the answer. If the problem requires you to execute commands in \texttt{R}, please include the code you used to get your answers. Please also include the \texttt{.R} file that contains your code. If you are not sure if work needs to be shown for a particular problem, please ask.
		\item Your homework should be submitted electronically on GitHub in \texttt{.pdf} form.
		\item This problem set is due before 23:59 on Sunday February 18, 2024. No late assignments will be accepted.
	%	\item Total available points for this homework is 80.
	\end{itemize}

	
	%	\vspace{.25cm}
	
%\noindent In this problem set, you will run several regressions and create an add variable plot (see the lecture slides) in \texttt{R} using the \texttt{incumbents\_subset.csv} dataset. Include all of your code.

	\vspace{.25cm}
%\section*{Question 1} %(20 points)}
%\vspace{.25cm}
\noindent We're interested in what types of international environmental agreements or policies people support (\href{https://www.pnas.org/content/110/34/13763}{Bechtel and Scheve 2013)}. So, we asked 8,500 individuals whether they support a given policy, and for each participant, we vary the (1) number of countries that participate in the international agreement and (2) sanctions for not following the agreement. \\

\noindent Load in the data labeled \texttt{climateSupport.RData} on GitHub, which contains an observational study of 8,500 observations.

\begin{itemize}
	\item
	Response variable: 
	\begin{itemize}
		\item \texttt{choice}: 1 if the individual agreed with the policy; 0 if the individual did not support the policy
	\end{itemize}
	\item
	Explanatory variables: 
	\begin{itemize}
		\item
		\texttt{countries}: Number of participating countries [20 of 192; 80 of 192; 160 of 192]
		\item
		\texttt{sanctions}: Sanctions for missing emission reduction targets [None, 5\%, 15\%, and 20\% of the monthly household costs given 2\% GDP growth]
		
	\end{itemize}
	
\end{itemize}

\newpage
\noindent Please answer the following questions:

\begin{enumerate}
	\item
	Remember, we are interested in predicting the likelihood of an individual supporting a policy based on the number of countries participating and the possible sanctions for non-compliance.
	\begin{enumerate}
		\item [] Fit an additive model. Provide the summary output, the global null hypothesis, and $p$-value. Please describe the results and provide a conclusion.
		%\item
		%How many iterations did it take to find the maximum likelihood estimates?
	\end{enumerate}
	\vspace{.8cm}
	\textbf{		The R script for additive model, summary output and p-value is:}
	\lstinputlisting[language=R, firstline=39, lastline=64]{PS2.R} 
	\vspace{.8cm}
	\begin{tabular}
		{ |p{4cm}||p{3cm}|p{3cm}|p{3cm}|p{3cm}  }
		\hline
		\multicolumn{4}{|c|}{Model Outputs:} \\
		\hline
		Coefficients&Estimate&Std. Error &z value &Pr(>|z|)\\
		\hline
		(Intercept)       &-0.08081    &0.05316  &-1.520  &0.12848\\
		countries80 of 192   &0.33636    &0.05380   &6.252 &4.05e-10 ***\\
		countries160 of 192  &0.64835    &0.05388  &12.033  &< 2e-16 ***\\ sanctionsNone       &-0.19186    &0.06216  &-3.086  &0.00203 **\\ sanctions15\%        &-0.32510    &0.06224  &-5.224 &1.76e-07 ***\\ sanctions20\%       & -0.49542    &0.06228  &-7.955 &1.79e-15 ***\\
		
		\hline
		
		\end{tabular}



\text {Signif. codes:  0 ‘***’ 0.001 ‘**’ 0.01 ‘*’ 0.05 ‘.’ 0.1 ‘ ’ 1 }
\newline {(Dispersion parameter for binomial family taken to be 1)}
\newline Null deviance: 11783  on 8499  degrees of freedom
\newline Residual deviance: 11568  on 8494  degrees of freedom

AIC: 11580

Number of Fisher Scoring iterations: 4

The p-values below the benchmark of 5\% show that both the explanatory variables \textbf{countries} and \textbf{sanctions} has signifcant impact on the response variable \textbf{choice}.


\vspace{.8cm}

	\item If any of the explanatory variables are significant in this model, then:
	\begin{enumerate}
		\item
		For the policy in which nearly all countries participate [160 of 192], how does increasing sanctions from 5\% to 15\% change the odds that an individual will support the policy? (Interpretation of a coefficient)
%		\item
%		For the policy in which very few countries participate [20 of 192], how does increasing sanctions from 5\% to 15\% change the odds that an individual will support the policy? (Interpretation of a coefficient)
\vspace{.5cm}
	
		\textbf{		The R script is:}
		\lstinputlisting[language=R, firstline=67, lastline=74]{PS2.R} 
		
		The probability of 15\% sanctions is estimated to be approx. 72\%.
		\newpage
		\item
		What is the estimated probability that an individual will support a policy if there are 80 of 192 countries participating with no sanctions? 
		
		\textbf{		The R script for probability estimation is:}
		\lstinputlisting[language=R, firstline=77, lastline=89]{PS2.R} 
		
		The probability of 80 countries out of 192 countries particpating with no sanctions is estimated to be 51.6\%.
		\vspace{.5cm}
		\item
		Would the answers to 2a and 2b potentially change if we included the interaction term in this model? Why? 
		\begin{itemize}
			\item Perform a test to see if including an interaction is appropriate.
		\end{itemize}
		\vspace{.5cm}
			\textbf{		The R script for interaction model is:}
		\lstinputlisting[language=R, firstline=92, lastline=98]{PS2.R} 
		
		Model 1: choice ~ countries + sanctions
		
		Model 2: choice ~ countries + sanctions + countries:sanctions
		
		Resid. Df Resid. Dev Df Deviance Pr(>Chi)
		\newline1      8494      11568                     
		\newline2      8488      11562       6        6.2928          0.3912
		\vspace{.5cm}
		
		The p-value above the benchmark of 5\% shows that the interaction term of \textbf{countries} and \textbf{sanctions} does not have a signifcant impact on the response variable \textbf{choice}.
		
		
	\end{enumerate}
	\end{enumerate}


\end{document}
